{\bf Intrusion Detection using Parzen-Windows on Provenance Graph Statistics}

### 2013-12-09 writeup/paper-final.tex
wrote related work

Existing intrusion detection systems (IDS) have used provenance data only to a limited extent.
??? person analyzes
basic statistics on provenance graphs (e.g., number of nodes, number of edges, ... TODO) in order to facilitate
easier manual intrusion detection by a human (CITE WORK). Their work, however, does not present a way of using provenance
graph to automatically determine intrusions. Other systems use provenance data to determine the scope of an
intrusion. For example, ??? Tariq use provenance graphs to build causality graphs of intrusions: once they have
determined an intrusion has occurred, they follow the causality graph backwards determine the original source of the attack, and 
they follow the causality graph forwards to determine all the objects tainted by the attack (CITE WORK). However, they also
do not present a way of using provenance data to automatically detect intrusions.

@@

In this paper, we present an intrusion detection approach using Parzen-Windows density estimation. ???? have used
Parzen-Windows for intrusion detection with a high degree of success in identifying various intrusion types (CITE WORK).
They justify using Parzen-Windows for
{\em novelty detection} when access to novel or abnormal patterns is difficult to obtain, as is the case with
intrusions. We borrow their Parzen-Windows approach to build models of normal behavior using 
various provenance graph statistics and graph centrality metrics, and we evaluate the success of this technique
on {\em user-to-local} (u2l) intrusions, intrusions that provide unauthorized access to local superuser (root) privileges.

>>>>

Existing intrusion detection systems (IDS) use provenance data only to a limited extent.
??? person analyzes
basic statistics on provenance graphs (e.g., number of nodes, number of edges, ... TODO) in order to facilitate
easier manual intrusion detection by a human (CITE WORK). Their work, however, does not present a way of using provenance
graph to automatically determine intrusions. Other systems use provenance data to determine the scope of an
intrusion. For example, Backtracker (CITE WORK) uses provenance graphs to build causality graphs of intrusions: once the
system has
determined an intrusion has occurred, the system follow the causality graph backwards determine the original source of the attack, and 
then it follow the causality graph forwards to determine all the objects tainted by the attack. However, the authors 
do not present a way of using provenance data to automatically detect intrusions.

@@

In this paper, we present an approach to use provenance data to automatically detect intrusions. Intrusion
detection can be framed as a {\em novelty detection} problem, in which one attempts to decide whether
an unknown test pattern is produced by an underlying distribution corresponding to a training set
of normal patterns (CITE WORK). However, ???authors note that in the case of intrusion detection, novel
or abnormal patterns are typically difficult to obtain, and as a result, they present a {\em Parzen-Window} approach
for non-parametric density estimation. Using this technique, they obtain a high degree of success
in identifying various intrusion types (CITE WORK). Because obtaining abnormal patterns is difficult, 
we borrow their Parzen-Window approach to build models of normal behavior using 
various provenance graph statistics and graph centrality metrics, and we evaluate the success of this technique
on {\em user-to-local} (u2l) intrusions, intrusions that provide unauthorized access to local superuser (root) privileges.
